Initial spatial conditions in simulation models: the missing leg of sensitivity analyses?

### 2017-07-21 paper.tex
Update on Overleaf.

In this paper, we suggest tackling sensitivity to spatial conditions by generating a variety of density grids to feed into simulation models at initialisation, and exploring the sensitivity of the model outcomes to the variation of meta-parameters (i.e. parameters used to generate initial spatial conditions). The purpose is two-fold: 1/ to test the robustness of simulation results to small variations of meta-parameters within a typical category of space (a monocentric case for example) and 2/ to study the non-trivial effects of typical categories of spatial distribution (monocentric vs. polycentric for example) on the results of a given model.

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Why spatial patterns are expected to impact social simulations}
\comment[MLT]{Je me demande si tout ce qui vient après jusqu'à la littérature review sur les attempts de prise en compte de l'effet de la configuration spatiale sur le comportement et les résultats des modèles ne serait pas mieux en intro? Ca ferait une longue intro mais ça éviterait de trop se répéter?}
An urban system consists of a large number of social agents that interact with each other at a microscopic level in a system-time that evolves irreversibly creating temporal and cumulative effects. Self-organization has been shown to be a central feature of urban systems~\citep{AllenSanglier1981,Portugali2000}, and emergent properties at qualitatively differentiated scales~\citep{AzizAlaouiBertelle2009} must be obtained through simulation methods~\citep{Wu2002,Batty2007}. In that frame, complexity is partially due to bifurcations \citep{Wilson1981} which are determinant in spatial systems~\citep{Wilson2002}.

@@

\paragraph{Previous considerations of urban form in simulation models}

%\comment[FL]{Je pense que cette partie est vraiment utile, et devrait être étoffée, en procédant peut être à une typologie plus systématique des "initial condition" qui nous intéressent. Il y a déjà une différence entre l'hétérogénéité de l'espace (densités réparties mono ou poly, etc) et les différentes façons de représenter l'espace (grilles carrées, hexagones). Par ailleurs aussi des discussions sur l'incertitude de l'information (est ce que par exemple les emplois downscalés au carreau INSEE par Pivano 2016 c'est vraiment fiable ou très ?) et la précision de la grille}

>>>>

An urban system consists of a large number of social agents that interact with each other at a microscopic level in a system-time that evolves irreversibly creating temporal and cumulative effects. Self-organization has been shown to be a central feature of urban systems~\citep{AllenSanglier1981,Portugali2000}, and emergent properties at qualitatively differentiated scales~\citep{AzizAlaouiBertelle2009} must be obtained through simulation methods~\citep{Wu2002,Batty2007}. In that frame, complexity is partially due to bifurcations \citep{Wilson1981} which are determinant in spatial systems~\citep{Wilson2002}.

@@

In this paper, we suggest tackling sensitivity to spatial conditions by generating a variety of density grids to feed into simulation models at initialisation, and exploring the sensitivity of the model outcomes to the variation of meta-parameters (i.e. parameters used to generate initial spatial conditions). The purpose is two-fold: 1/ to test the robustness of simulation results to small variations of meta-parameters within a typical category of space (a monocentric case for example\comment[FL]{je ne trouve pas cela facile à comprendre car on n'a pas encore parlé des différentes formes de villes}) and 2/ to study the non-trivial effects of typical categories of spatial distribution (monocentric vs. polycentric for example) on the results of a given model.

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Why spatial patterns are expected to impact social simulations}

\comment[MLT]{Je me demande si tout ce qui vient après jusqu'à la littérature review sur les attempts de prise en compte de l'effet de la configuration spatiale sur le comportement et les résultats des modèles ne serait pas mieux en intro? Ca ferait une longue intro mais ça éviterait de trop se répéter?}

\paragraph{Previous considerations of urban form in simulation models}
\begin{itemize}
\item spatial context
\begin{itemize}
\item extent
\item objects
\item shape
\end{itemize}

\item spatial "encoding"
\begin{itemize}
\item heterogeneity
\item algorithm of desagregation
\item uncertainty
\end{itemize}

\item spatial effects

\begin{itemize}
\item topology
\item interaction (??)
\end{itemize}
\item spatial outputs
\begin{itemize}
\item agregation
\item use of models
\end{itemize}
\end{itemize}
%\comment[FL]{Je pense que cette partie est vraiment utile, et devrait être étoffée, en procédant peut être à une typologie plus systématique des "initial condition" qui nous intéressent. Il y a déjà une différence entre l'hétérogénéité de l'espace (densités réparties mono ou poly, etc) et les différentes façons de représenter l'espace (grilles carrées, hexagones). Par ailleurs aussi des discussions sur l'incertitude de l'information (est ce que par exemple les emplois downscalés au carreau INSEE par Pivano 2016 c'est vraiment fiable ou très ?) et la précision de la grille}
